Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-04-30 overleaf/introduction.tex
Made a minor edit.
Edits to 1-5.

We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account
control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.

@@

\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code. %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}

@@

Our most important contribution is the implementation and evaluation of the compiler framework. We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others. Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol. Results form BMR are similar.

>>>>

We use the MPC Source~\cite{Ishaq:2019} intermediate representation and emphasize optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., our front end IMP Source, other) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account
control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.

@@

\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP Source language, its syntax and semantic restrictions, (b) translation from IMP Source into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code. %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}

@@

Our most important contribution is the implementation and evaluation of the compiler framework. We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others. \ana{Summarize results.}

### 2022-04-30 overleaf/vectorization.tex
Edits to 1-5.

Phase 2 proceeds from the inner-most towards the outer-most loop. For each loop it anchors dependence cycles (closures) 
around pseudo PHI nodes then removes vectorization from the dimension of that loop. There are two important points in this 
phase. First, it may break a loop into smaller loops which would discover opportunities 
for vectorization in intermediate statements in the loop. Second, it handles write arrays. 
It creates opportunities for vectorization in the presence of write arrays, even though 
Cytron's SSA adds a back-edge to the array PHI-node, thus killing vectorization of statements
that read and write that array.

The excerpts in {\color{red} red} color in the pseudo code below highlight the extension 
with array writes. We advise the reader to omit the extension and assume just read-only
arrays. We explain the extension in \secref{sec:array_writes} details our algorithm. 
(As many of our benchmarks include array writes, it plays an important role.)

>>>>

Phase 2 proceeds from the inner-most towards the outer-most loop. For each loop it anchors dependence cycles (closures) 
around pseudo PHI nodes then removes vectorization from the dimension of that loop. There are two important points in this 
phase. First, it may break a loop into smaller loops which would discover opportunities 
for vectorization in intermediate statements in the loop. Second, it handles writes arrays. 
It creates opportunities for vectorization in the presence of write arrays, even though 
Cytron's SSA adds a back-edge to the array PHI-node, thus killing vectorization of statements
that read and write that array.

The excerpts in {\color{red} red} color in the pseudo code below highlight the extension 
with array writes. We advise the reader to omit the extension for now and consider just read-only
arrays. We explain the extension in \secref{sec:array_writes}. 
(As many of our benchmarks include write arrays, it plays an important role.)
